### 2023-01-09 Manuscript/manuscript.tex
update

% ****** Start of file apssamp.tex ******
%
%   This file is part of the APS files in the REVTeX 4.2 distribution.
%   Version 4.2a of REVTeX, December 2014
%
%   Copyright (c) 2014 The American Physical Society.
%
%   See the REVTeX 4 README file for restrictions and more information.
%
% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
% as well as the rest of the prerequisites for REVTeX 4.2
%
% See the REVTeX 4 README file
% It also requires running BibTeX. The commands are as follows:
%
%  1)  latex apssamp.tex
%  2)  bibtex apssamp
%  3)  latex apssamp.tex
%  4)  latex apssamp.tex
%
\documentclass[%
% reprint,
%superscriptaddress,
%groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
preprint,
%preprintnumbers,
%nofootinbib,
%nobibnotes,
%bibnotes,
 amsmath,amssymb,
 aps,
%pra,
%prb,
%rmp,
%prstab,
%prstper,
%floatfix,
]{revtex4-2}

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%\usepackage{hyperref}% add hypertext capabilities
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

%\usepackage[showframe,%Uncomment any one of the following lines to test 
%%scale=0.7, marginratio={1:1, 2:3}, ignoreall,% default settings
%%text={7in,10in},centering,
%%margin=1.5in,
%%total={6.5in,8.75in}, top=1.2in, left=0.9in, includefoot,
%%height=10in,a5paper,hmargin={3cm,0.8in},
%]{geometry}

\begin{document}

\preprint{}

\title{Heart stroke-volume variability in a murine model for heart failure with reduced ejection fraction}% Force line breaks with \\
%\thanks{A footnote to the article title}%

\author{Gemma Fernández-Mendoza}
 \email{gemmag.fernandez@gmail.com}
\affiliation{%
 Departamento de Física, Escuela Superior de Física y Matemáticas, Instituto Politécnico Nacional, 07738 Ciudad de México, México
}%

\author{Moisés Santillán}
 \homepage{http://moises-santillan.github.io}
 \email{msantillan@cinvestav.mx}
\affiliation{
 Centro de Investigación y de Estudios Avanzados, Unidad Monterrey, 66628 Apodaca NL, México
}%

\date{\today}% It is always \today, today,
             %  but any date may be explicitly specified

\begin{abstract}
Write an abstract
\end{abstract}

%\keywords{Suggested keywords}%Use showkeys class option if keyword
                              %display desired
\maketitle

%\tableofcontents

\section{\label{sec:intro}Introduction}

\section{Materials and Methods}

\section{Results}

\subsection{Experimental Results}

\subsection{Variability analysis}

Heart rate variability has been measured using a variety of techniques. The majority of them are based on the notion of signal stationarity. However, the heart rate's inherently nonstationary nature---which undergoes continuous physiological change to adapt to outside stimuli---presents a significant challenge that could lead to inaccurate results \citep{Marwan_2007}. Although a number of signal preprocessing methodologies have been suggested to address these problems, nonlinear analysis-based strategies are frequently used and seem to produce reliable result \citep{Marwan_2002, Aubert_2003, Marwan_2007, Giuliani_1998, Rajendra_Acharya_2006, Webber_1994, Henriques_2020}. One of them that is used in different scientific domains is the Poincairé plot \citep{Hoshi_2016, Webber_1994}.

\subsection{Mathematical Model}

\begin{figure}
\includegraphics[width=4in]{model.pdf}
\caption{Scematic representation of the model}
\label{fig:model}
\end{figure}

\section{Concluding Remarks}

\begin{acknowledgments}
We wish to acknowledge the support of the author community in using
REV\TeX{}, offering suggestions and encouragement, testing new versions,
\dots.
\end{acknowledgments}

\bibliography{manuscript}% Produces the bibliography via BibTeX.

\end{document}
%
% ****** End of file apssamp.tex ******